### 2020-04-30 submit/cover_letter.tex
cover letter

\documentclass[12pt,letterpaper]{article}
\input{plrletter-commands}
\usepackage[hidelinks]{hyperref}

\signature{
    Peter Ralph, on behalf of Kevin Thornton and Jerome Kelleher.}
\name{Peter Ralph}
\address{Peter Ralph\\ Institute for Ecology and Evolution\\ and Departments of Mathematics and Biology,\\ University of Oregon\\
plr@uoregon.edu}
\telephone{(707) 502-5854}
\email{plr@uoregon.edu}

\begin{document}

\pagestyle{plain}

\noindent

\opening{To the editor(s) --}
\hspace{\stretch{1}}
\today

We are writing with great joy and the final version of our accepted manuscript,
``Efficiently summarizing relationships in large samples: a general duality between statistics of genealogies and genomes''.
We have addressed both of Reviewer 2's final suggestions,
by adding an explanatory note to Figure 1's caption, and deleting the word "significant",
which didn't add anything to the sentence it was in.
No other changes have been made,
other than moving the \LaTeX over to the Genetics template.

We have not added section titles for ``Materials and Methods'' or ``Results'',
since we don't feel those are appropriate for our theoretically-oriented paper.
We hope this is OK with you,
as we see other fine Genetics papers have not included these section headings
(e.g., Jain \& Stephan 2017, \url{https://www.genetics.org/content/206/1/389}).
But, if you'd prefer we'll gladly put them in --- just let us know.

\closing{Thanks very much,}

\end{document}